---
title: What Do People Google During Debates?
author: "Janalee Thompson, Scott Kelleher, Tulsi Ram Gompo"
date: "11/15/2016"
output: html_document
---

**##2016 Presidential Debate: Content Breakdown**

  The 2016 presidential debates unfolded slowly as November approached. We watched
candidates battle to the bittersweet end trying to convince the American people that
they were the better-suited candidate to lead our country. According to CNN.com, over
84 million viewers tuned in to watch Hillary Clinton and Donald Trump debate. There
were three debates in total spanning over a couple months. Some would consider these
debates amusing (or scary) to say the least, such that terminology never before used in
political discussions was broadcasted live for United States citizens and other
countries to see in all its' glory. "Piggy" and "deplorable" were just a couple
words that were projected during the debates that caught the attention of some viewers.
So what is it that we all do when we want to know more about something? We just "Google
it" of course. 

**What were we thinking?**

  As each debate unfolded, we wondered how many other people (like us) were using
Google to search for termininology used in the debates- like the second amendment for
example. Thus, we generated an analysis utilizing debate content and Google search
terms. We started checking out Google Trends and tried trending some search terms
ourselves. Google Trends is available to the general public with the purpose of
displaying global search trends in Google. Google Trends adjusts Google searches by
dividing each individual data point by total searches, geography and time range. In
other words, this displays the term that was searched in relation to global
search-volume. Trends in the data can be graphed by country or city. 
  
  Terms such as “second amendment” and “Isis” were some of the first terms we tried
that we knew would probably prompt some pretty drastic search trends, which they did.
Then, after using Google Trends to search what was intended to be a few words, we ended
up searching a copious amount of terms (related and unrelated to the debates)! It was
time to get down to business and locate the debate scripts and take some terms used by
our candidates for a spin in Google Trends. 

  We read through each debate script and noted some terminology that stood out.
Displayed below are some terms used by our now President-Elect Donald Trump during the
second debate. The terms `Melania`, `e-mail scandal`, `big league`, `nasty woman`, and
`benghazi' were searched using Google during and immediately following the second
debate. "Bigly" was misinterpreted as the Donald Trump phrase, "Big League." As a
result of this confusion, Google Trends reported that "Bigly" had become the top
trending search on all of Google. Check out the Tweet
[here](http://www.aol.com/article/news/2016/10/19/1-word-most-searched-term-during-pres
idential-debate/21587735/). 

**Five Phrases that had particularly interesting Google search trends with respect to
their debate**
```{r echo = FALSE, message = FALSE}
##Loading libraries
library(rvest)
#library(tidyverse)
library(stringr)
library(tidytext)
library(gtrendsR)
library(tokenizers)
library(dplyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(sentimentr)
library(lubridate)
library(ggplot2)
library(readr)
library(gtrendsR)
library(xml2)
library(shiny)
library(devtools)
ls("package:gtrendsR")

#library(shinydashboard)
library(RTextTools)
library(googleVis)
library(sentimentr)
library(sentiment)
source("functions/classify_emotion.R")

library(RTextTools)
library(googleVis)
library(DT)
library(choroplethr)
library(choroplethrMaps)


usr <- ("535rprogram@gmail.com")
  psw <- ("groupproject")
  ch <- gconnect(usr, psw)
```

```{r, echo = FALSE, message = FALSE}
  Speaker <- "Donald Trump"  # enter Donald_Trump or Hilary_Clinton
  Debate = "Second_Debate"   # enter First_Debate or Second_Debate or Third_Debate
  google_these_words = c("Melania", "email scandal", "bigly", "nasty woman", "benghazi")  # enter up to five phrases to google in gtrends and results
  # enter one phrase you would like to see broken down by state

```

  Additionally, we were interested in specific terminology searched in Google at the
state level. Specifically, which states were Googling more particular terms than
others? The following graphic is categorized by state using Google Trends search term
data. This time we used the search term "Clinton Foundation" and attained the following
result. 
```{r, echo = FALSE, message = FALSE}
google_states = c("Clinton Foundation")

```

**Emotions**










plots to include:
gtrends with time: 5 interesting phrases
plots by state pick phrase to plot by state
emotions






```{r, echo=FALSE,message=FALSE}
library(rvest)
library(tidyverse)
library(stringr)
library(tidytext)
library(gtrendsR)
library(tokenizers)
library(dplyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(lubridate)
library(ggplot2)
library(readr)
library(gtrendsR)
library(gtrendsR)
library(RTextTools)


##```{r,echo=FALSE, message=FALSE }
text_debate1 <- read_html("http://www.presidency.ucsb.edu/ws/index.php?pid=118971")  # load the first debate page
text_debate2 <- read_html("http://www.presidency.ucsb.edu/ws/index.php?pid=119038")   # load the second debate page
text_debate3 <- read_html("http://www.presidency.ucsb.edu/ws/index.php?pid=119039")  # load the third debate page 
    
text_debate1 <- html_nodes(text_debate1, ".displaytext") %>%   # isloate the text
html_text()    # get the text
  
text_debate2 <- html_nodes(text_debate2, ".displaytext") %>% # isloate the text
html_text() # get the text
      
text_debate3 <- html_nodes(text_debate3, ".displaytext") %>% # isloate the text
html_text() # get the text

w <- c(text_debate1, text_debate2, text_debate3)
```



##``` {r, echo=FALSE,message=FALSE}
 #dates of debate
db1 <- ymd_hms(160927020000, tz = "UTC")
db2 <- ymd_hms(161009020000, tz = "UTC")
db3 <-ymd_hms(161029020000, tz = "UTC")
date_time <-c(db1, db2, db3)
```


