---
title: What Do People Google During Debates?
author: "Janalee Thompson, Scott Kelleher, Tulsi Ram Gompo"
date: "11/15/2016"
output: html_document
---

**Introduction**
The 2016 presidential debates unfolded slowly as November approached. We watched
candidates battle to the bittersweet end trying to convince the American people that
they were the better-suited candidate to lead our country. According to CNN.com, over 84
million people tuned in to watch Hillary Clinton and Donald Trump debate. The three
debates were amusing to say the least, such that terminology never before used in
political discussions was used to deflate the opponent. Though after each debate, we
were left wondering how this debate was broken down according to Google. Thus, we
generated an analysis utilizing debate content and Google search terms. 
	  We started checking out Google Trends and tried out some search terms. Google Trends
is available to the general public with the purpose of displaying trends in global
Google searches. Google Trends adjusts Google searches by dividing each individual data
point by total searches, geography and time range. Trends in the data are graphed by
country or city. Terms such as “second amendment” and “Isis” were some of the first
terms we tried that we knew would probably prompt some pretty drastic search trends. We read through each debate script and found some terminology that stood out. 

**5 Phrases that had particularly interesting search trends with respect to their debate**
```{r echo = FALSE, message = FALSE}
```



gtrends with time: 5 interesting phrases

plots by state pick phrase to plot by state

emotions






```{r, echo=FALSE,message=FALSE}
library(rvest)
library(tidyverse)
library(stringr)
library(tidytext)
library(gtrendsR)
library(tokenizers)
library(dplyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(lubridate)
library(ggplot2)
library(readr)
library(gtrendsR)
library(gtrendsR)
library(RTextTools)


##```{r,echo=FALSE, message=FALSE }
text_debate1 <- read_html("http://www.presidency.ucsb.edu/ws/index.php?pid=118971")  # load the first debate page
text_debate2 <- read_html("http://www.presidency.ucsb.edu/ws/index.php?pid=119038")   # load the second debate page
text_debate3 <- read_html("http://www.presidency.ucsb.edu/ws/index.php?pid=119039")  # load the third debate page 
    
text_debate1 <- html_nodes(text_debate1, ".displaytext") %>%   # isloate the text
html_text()    # get the text
  
text_debate2 <- html_nodes(text_debate2, ".displaytext") %>% # isloate the text
html_text() # get the text
      
text_debate3 <- html_nodes(text_debate3, ".displaytext") %>% # isloate the text
html_text() # get the text

w <- c(text_debate1, text_debate2, text_debate3)
```



##``` {r, echo=FALSE,message=FALSE}
 #dates of debate
db1 <- ymd_hms(160927020000, tz = "UTC")
db2 <- ymd_hms(161009020000, tz = "UTC")
db3 <-ymd_hms(161029020000, tz = "UTC")
date_time <-c(db1, db2, db3)
```


