---
title: "What is the trend of vocabulary use among the presidential candidates during
  the debates in 2016 election ?"
author: Janalee Thompson, Scott Kelleher,Tulsi Ram Gompo
date: "11/16/2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```




# Overview of the Presentation





## 



- **Research Question**


- **Introduction**


- **Methods**


- **Results**


- **Conclusions**
 
## 

**Research Question**



Do the words spoken  by the presidental candidates at the debates matter a lot?


##

**Introduction**

The Presidential debates are most popular program in US and every body cares about what each presidential candidate said during the debates. While the presidential candiates are firing up with new words,the viewers either applauded them or booed widely.For example, the candidates Donald Trump used very strong words to address the opposition leader Hilary Clinton. Later, his words like “deplorable”, “wall” "nasty women " went viral and there was spike in search of those words in internet.The cliches and the hackneyed expresion not only creates boredome to viewer but also judge the quality of speech by debators .So, we are curious which president used what type of words and we decided to isolate top ten words that each president used in debates except the stop words.




## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure, message = FALSE, warning = FALSE}
library(plotly)
plot_ly(z = ~volcano)
```

